

###DATADELAY

```{r}
library(readr)
library(tidyverse)
library(datadelay)
library(epiparameter)
```

Loading outbreak scenario data from one of the workshop's groups

```{r}
Linelist_group3 <- read_csv("Data/Linelist_group3.csv")
head(Linelist_group3)
```

Formatting data as daily incidence of cases and deaths

```{r}
daily_cases_deaths <- Linelist_group3[,c(6,8)]
names(daily_cases_deaths)[1:2]=c("date","deaths")
daily_cases_deaths <- daily_cases_deaths %>% mutate(cases = rep(1, nrow(daily_cases_deaths)), .after = "date")
daily_cases_deaths$deaths <- ifelse(is.na(daily_cases_deaths$deaths)==T, 0, 1)
daily_cases_deaths <- aggregate(cbind(cases, deaths) ~ date, data = daily_cases_deaths, FUN = sum)

head(daily_cases_deaths)
```




```{r}
library(incidence2)
daily_cases_deaths <- incidence2::incidence(Linelist_group3,c("onset_date","death_date")) |> complete_dates()
daily_cases_deaths <- pivot_wider(daily_cases_deaths, names_from = count_variable, values_from = count)
names(daily_cases_deaths)[2:3]=c("deaths","cases")
daily_cases_deaths <- daily_cases_deaths[c(1,3,2)]
daily_cases_deaths$date_index <- as.Date(daily_cases_deaths$date_index)
daily_cases_deaths <- as.data.frame(daily_cases_deaths)
```

Static CFR- naive estimate

```{r}
n_static_cfr <- estimate_static(daily_cases_deaths, correct_for_delays = FALSE) 
n_static_cfr
```

Static CFR- adjusted estimate, taking ebola onset to death dist from `epiparameter`, *getting ERROR*:

```{r}
onset_to_death_ebola <- epiparameter::epidist_db(
  disease = "Ebola Virus Disease",
  epi_dist = "onset_to_death",
  author = "WHO_Ebola_Response_Team")

c_static_cfr <- estimate_static(daily_cases_deaths, epi_dist = onset_to_death_ebola, correct_for_delays = T)
```

Naive time-varying CFR
*Now it's necessary to add an epidist object even if you are not correcting for delays, otherwise you get an ERROR- this should be highlighted in the documentation?*

```{r}
n_rolling_CFR <- estimate_time_varying(daily_cases_deaths, epi_dist = onset_to_death_ebola, correct_for_delays = FALSE)
head(n_rolling_CFR)
```

Corrected time-varying CFR 
*No error*

```{r}
c_rolling_CFR <- estimate_time_varying(daily_cases_deaths, epi_dist = onset_to_death_ebola)

```

Plotting CFR, *there used to be a function `plot_data_and_cfr()` to plot both naive and corrected CFR that generated some nice plots, but now it's been removed?*

```{r}
plot_time_varying(n_rolling_CFR)
```

```{r}
plot_time_varying(c_rolling_CFR)
```

*The README on `datadelay` hasn't been updated so the old functions are still there, which is confusing for users*

###DATADELAY WITH DATA CREATED FOR REPEX 

```{r}
#Creating linelist
daily_cases_deaths <- data.frame(case_id = 1:100) %>% 
  mutate(date = sample(seq(as.Date("2023-01-01"), by = "day", length.out = n()))) %>%
  mutate(cases = sample(1:10, replace = T, n())) %>%
  mutate(deaths = sample(0:3, replace = T, n()))

#Introducing missing values
daily_cases_deaths <- daily_cases_deaths[-sample(nrow(daily_cases_deaths), 10), ]

#Using `estimate_static` with correct=T, which results in an error:

library(epiparameter)
library(datadelay)

onset_to_death_ebola <- epiparameter::epidist_db(
  disease = "Ebola Virus Disease",
  epi_dist = "onset_to_death",
  author = "WHO_Ebola_Response_Team")

c_static_cfr <- estimate_static(daily_cases_deaths, epi_dist = onset_to_death_ebola, correct_for_delays = T)
```

### Repex cleanepi

```{r}
linelist <- data.frame(case_id = 1:100) %>%
  mutate(
    onset_date = seq(as.Date("2023-01-01"), by = "day", length.out = n())
  ) %>%
  mutate(death_date = seq(as.Date("2023-02-01"), by = "day", length.out = n())) %>%
  mutate(age = sample(seq(0:90), replace = T, n())) %>%
  mutate(gender = sample(0:1, replace = T, n()))

```


```{r}
library(cleanepi)
linelist_clean <- cleanepi(linelist, remove.duplicates = T, duplicates.from = c("case_id","onset_date","death_date","age","gender"))
```

```{r}

linelist3_clean <- cleanepi(Linelist_group3, remove.duplicates = T, duplicates.from = c("case_name","ct_value","age","hospitalisation_date","date_first_contact","report_date","occupation","case_type","gender","onset_date","death_date","date_last_contact","island_group"))

```





